### 2016-02-15 intro/intro.Rmd
updates to intro
re-organization of repo

```{r}
spike.f=function(x) (0.75*exp(-500*(x-0.23)^2)+1.5*exp(-2000*(x-0.33)^2)+3*exp(-8000*(x-0.47)^2)+2.25*exp(-16000*(x-0.69)^2)+0.5*exp(-32000*(x-0.83)^2))
n=1024
t=1:n/n
mu.s=spike.f(t)

library(wavethresh)
library(EbayesThresh)
library(caTools)
source(file.path("../Rcode/bayesmooth.R"))
source(file.path("../Rcode/ti_thresh.R"))

mse=function(x,y) mean((x-y)^2)
l2norm=function(x) sum(x^2)
mise=function(x,y,r) 10000*mean(apply(x-rep(1,r)%o%y,1,l2norm)/l2norm(y))
sig.est.func=function(x,n) sqrt(2/(3*(n-2))*sum((1/2*x[1:(n-2)]-x[2:(n-1)]+1/2*x[3:n])^2))

##Default R wavelet method, written into a single function
waveti.u=function (x, filter.number = 10, family = "DaubLeAsymm", min.level = 3, noise.level){
    TT <- length(x)
    thresh <- noise.level * sqrt(2 * log(TT))    
    x.w <- wd(x, filter.number, family, type = "station")
    x.w.t <- threshold(x.w, levels = (min.level):(x.w$nlevels - 1), policy = "manual", value = thresh, type = "hard")
    x.w.t.r <- AvBasis(convert(x.w.t))
    return(x.w.t.r)
}

waveti.ebayes=function (x, filter.number = 10, family = "DaubLeAsymm", min.level = 3, noise.level){
    n=length(x)
    J=log2(n)
    x.w <- wd(x, filter.number, family, type = "station")
    for(j in min.level:(J-1)){
      x.pm = ebayesthresh(accessD(x.w,j),sdev=noise.level)
      x.w = putD(x.w,j,x.pm)
    }
    mu.est=AvBasis(convert(x.w))
    return(mu.est)
}

@@

mu.est=apply(X.s,1,bayesmooth)
mu.est.tivar.ash=apply(X.s,1,ti.thresh,method="bayesm")
mu.est.tivar.mad=apply(X.s,1,ti.thresh,method="rmad")

@@

##Bayesmooth
mise(t(mu.est),mu.t,10)
##TI thresholding with variance estimated from bayesmooth
mise(t(mu.est.tivar.ash),mu.t,10)
##TI thresholding with variance estimated from running MAD
mise(t(mu.est.tivar.mad),mu.t,10)
##TI thresholding with constant variance (estimated)
mise(mu.est.ti,mu.t,10)
##EBayes with constant variance (estimated)
mise(mu.est.ti.ebayes,mu.t,10)
```

@@

plot(mu.t,type='l')
lines(mu.est[,1],col=2)
lines(mu.est.tivar.mad[,1],col=3)
lines(mu.est.ti[1,],col=4)
lines(mu.est.ti.ebayes[1,],col=6)
```

@@

mu.est=apply(X.s,1,bayesmooth)
mu.est.tivar.ash=apply(X.s,1,ti.thresh,method="bayesm")
mu.est.tivar.mad=apply(X.s,1,ti.thresh,method="rmad")

@@

##Bayesmooth
mise(t(mu.est),mu.t,10)
##TI thresholding with variance estimated from bayesmooth
mise(t(mu.est.tivar.ash),mu.t,10)
##TI thresholding with variance estimated from running MAD
mise(t(mu.est.tivar.mad),mu.t,10)
##TI thresholding with constant variance (estimated)
mise(mu.est.ti,mu.t,10)
##EBayes with constant variance (estimated)
mise(mu.est.ti.ebayes,mu.t,10)

@@

```{r fig.width=12, fig.height=12}
plot(mu.t,type='l')
lines(mu.est[,1],col=2)
lines(mu.est.tivar.mad[,1],col=3)
lines(mu.est.ti[1,],col=4)
lines(mu.est.ti.ebayes[1,],col=6)

@@

pos = c(.1, .13, .15, .23, .25, .40, .44, .65, .76, .78, .81)
hgt = c(4, 5, 3, 4, 5, 4.2, 2.1, 4.3, 3.1, 5.1, 4.2)
wth = c(.005, .005, .006, .01, .01, .03, .01, .01, .005, .008, .005)
mu.b = rep(0,n)
for(j in 1:length(pos)){
  mu.b = mu.b + hgt[j]/(( 1 + (abs(t - pos[j])/wth[j]))^4)
}
dop.f=function(x) sqrt(x*(1-x))*sin((2*pi*1.05)/(x+0.05))
mu.dop=dop.f(t)

@@

var.est.1=bayesmooth(X.s.1,v.est=TRUE)
var.est.2=bayesmooth(X.s.2,v.est=TRUE)
var.est.3=bayesmooth(X.s.3,v.est=TRUE)
var.est.4=bayesmooth(X.s.4,v.est=TRUE)

@@

```{r fig.width=12, fig.height=12}
par(mfrow=c(2,2))
plot(sigma.t.1^2,type='l')
lines(var.est.1,col=2)
plot(sigma.t.2^2,type='l')
lines(var.est.2,col=2)
plot(sigma.t.3^2,type='l')
lines(var.est.3,col=2)
plot(sigma.t.4^2,type='l')
lines(var.est.4,col=2)

@@

Now we take a look at how this wavelet-based approach handles unevenly spaced data. Wavelet methods are not optimized for unevenly spaced, but still works.

@@

##Bayesmooth
mise(t(mu.est),mu.t,10)
##TI thresholding with variance estimated from bayesmooth
mise(t(mu.est.tivar.ash),mu.t,10)
##TI thresholding with variance estimated from running MAD
mise(t(mu.est.tivar.mad),mu.t,10)
##TI thresholding with constant variance (estimated)
mise(mu.est.ti,mu.t,10)
##EBayes with constant variance (estimated)
mise(mu.est.ti.ebayes,mu.t,10)

```{r fig.width=12, fig.height=12}
par(mfrow=c(1,1))
plot(t,mu.t,type='l')
lines(t,mu.est[,1],col=2)
lines(t,mu.est.tivar.mad[,1],col=3)
lines(t,mu.est.ti[1,],col=4)
lines(t,mu.est.ti.ebayes[1,],col=6)
```

mu.est=apply(X.s,1,bayesmooth)
mu.est.tivar.ash=apply(X.s,1,ti.thresh,method="bayesm")
mu.est.tivar.mad=apply(X.s,1,ti.thresh,method="rmad")

##Bayesmooth
mise(t(mu.est),mu.t,10)
##TI thresholding with variance estimated from bayesmooth
mise(t(mu.est.tivar.ash),mu.t,10)
##TI thresholding with variance estimated from running MAD
mise(t(mu.est.tivar.mad),mu.t,10)
##TI thresholding with constant variance (estimated)
mise(mu.est.ti,mu.t,10)
##EBayes with constant variance (estimated)
mise(mu.est.ti.ebayes,mu.t,10)

@@

@@

  mu.est=bayesmooth(y.data)
  mu.est=mu.est[1:200]
  var.est=bayesmooth(y.data,v.est=TRUE)
  var.est=var.est[1:200]

@@

plot(x,mu.t,type='l')
lines(x,mu.est[1:200],col=2)

plot(x,sigma.t^2,type='l')
lines(x,var.est[1:200],col=2)

@@

```{r}
data=read.csv("../gaus_data/treas_bill/treas_bill.csv",header=FALSE)
data=as.numeric(data[,1])

@@

y.est=bayesmooth(y.final)
y.est.var=bayesmooth(y.final,v.est=TRUE)

@@

```{r fig.width=12, fig.height=12}
par(mfrow=c(2,2))
plot(data,type='l')

plot(x,y)
lines(x.mod,y.est[1:(length(y.mod))],col=2)

plot(x.mod,y.est[1:(length(y.mod))],type='l',ylim=c(-0.3,0.3))

plot(x.mod,sqrt(y.est.var[1:(length(y.mod))]),type='l')
````

@@

  mu.est<-bayesmooth(y.final)
  mu.est=mu.est[1:n]
  mse.hsm.c[i]=mse(mu.est,mu.t)

@@

  mu.est<-bayesmooth(y.final)
  mu.est=mu.est[1:n]
  mse.hsm.n[i]=mse(mu.est,mu.t)

@@

y.est=bayesmooth(y.final)
y.est=y.est[1:length(y)]
y.var.est=bayesmooth(y.final,v.est=TRUE)
y.var.est=y.var.est[1:length(y)]

@@

plot(x,y,ylim=c(-150,100))
lines(x,y.est,col=2)
plot(x,y.var.est,type='l')

@@

plot(x,y,ylim=c(-1,5))
lines(x,y.est,col=2)
plot(x,y.var.est,type='l')

>>>>

```{r}
#create mean function
spike.f=function(x) (0.75*exp(-500*(x-0.23)^2)+1.5*exp(-2000*(x-0.33)^2)+3*exp(-8000*(x-0.47)^2)+2.25*exp(-16000*(x-0.69)^2)+0.5*exp(-32000*(x-0.83)^2))
n=1024
t=1:n/n
mu.s=spike.f(t)

#load packages
library(wavethresh)
library(EbayesThresh)
library(caTools)
library(smash)

#define functions to be used later
mse=function(x,y) mean((x-y)^2)
l2norm=function(x) sum(x^2)
mise=function(x,y,r) 10000*mean(apply(x-rep(1,r)%o%y,1,l2norm)/l2norm(y))
sig.est.func=function(x,n) sqrt(2/(3*(n-2))*sum((1/2*x[1:(n-2)]-x[2:(n-1)]+1/2*x[3:n])^2))

##Default R wavelet thresholding method, written into a single function
waveti.u=function (x, filter.number = 10, family = "DaubLeAsymm", min.level = 3, noise.level){
    TT = length(x)
    thresh = noise.level * sqrt(2 * log(TT))    
    x.w = wd(x, filter.number, family, type = "station")
    x.w.t = threshold(x.w, levels = (min.level):(x.w$nlevels - 1), policy = "manual", value = thresh, type = "hard")
    x.w.t.r = AvBasis(convert(x.w.t))
    return(x.w.t.r)
}

##Similarly for Ebayesthresh
waveti.ebayes=function (x, filter.number = 10, family = "DaubLeAsymm", min.level = 3, noise.level){
    n=length(x)
    J=log2(n)
    x.w = wd(x, filter.number, family, type = "station")
    for(j in min.level:(J-1)){
      x.pm = ebayesthresh(accessD(x.w,j),sdev=noise.level)
      x.w = putD(x.w,j,x.pm)
    }
    mu.est=AvBasis(convert(x.w))
    return(mu.est)
}

@@

mu.est=apply(X.s,1,ashsmooth.gaus)
mu.est.tivar.ash=apply(X.s,1,ti.thresh,method="smash")
mu.est.tivar.mad=apply(X.s,1,ti.thresh,method="rmad")

@@

##SMASH
mise(t(mu.est),mu.t,10)
##TI thresholding with variance estimated from smash
mise(t(mu.est.tivar.ash),mu.t,10)
##TI thresholding with variance estimated from running MAD
mise(t(mu.est.tivar.mad),mu.t,10)
##TI thresholding with constant variance (estimated)
mise(mu.est.ti,mu.t,10)
##EBayes with constant variance (estimated)
mise(mu.est.ti.ebayes,mu.t,10)
```

@@

plot(mu.t, xlab = "", ylab = "",type='l')
lines(mu.est[,1],col=2)
lines(mu.est.tivar.mad[,1],col=3)
lines(mu.est.ti[1,],col=4)
lines(mu.est.ti.ebayes[1,],col=6)
legend("topright", legend = c("smash", "ti_rmad", "ti_homo", "ebayes_homo"), col = c(2, 3, 4, 6))
```

@@

mu.est=apply(X.s,1,ashsmooth.gaus)
mu.est.tivar.ash=apply(X.s,1,ti.thresh,method="smash", ashparam = list(nullcheck = FALSE))
mu.est.tivar.mad=apply(X.s,1,ti.thresh,method="rmad")

@@

##smash
mise(t(mu.est),mu.t,10)
##TI thresholding with variance estimated from smash
mise(t(mu.est.tivar.ash),mu.t,10)
##TI thresholding with variance estimated from running MAD
mise(t(mu.est.tivar.mad),mu.t,10)
##TI thresholding with constant variance (estimated)
mise(mu.est.ti,mu.t,10)
##EBayes with constant variance (estimated)
mise(mu.est.ti.ebayes,mu.t,10)

@@

```{r fig.width=12, fig.height=12}
plot(mu.t,xlab = "", ylab = "", type='l')
lines(mu.est[,1],col=2)
lines(mu.est.tivar.mad[,1],col=3)
lines(mu.est.ti[1,],col=4)
lines(mu.est.ti.ebayes[1,],col=6)
legend("topright", legend = c("smash", "ti_rmad", "ti_homo", "ebayes_homo"), col = c(2, 3, 4, 6))

@@

#define test functions
pos = c(.1, .13, .15, .23, .25, .40, .44, .65, .76, .78, .81)
hgt = c(4, 5, 3, 4, 5, 4.2, 2.1, 4.3, 3.1, 5.1, 4.2)
wth = c(.005, .005, .006, .01, .01, .03, .01, .01, .005, .008, .005)
mu.b = rep(0,n)
for(j in 1:length(pos)){
  mu.b = mu.b + hgt[j]/(( 1 + (abs(t - pos[j])/wth[j]))^4)
}
dop.f=function(x) sqrt(x*(1-x))*sin((2*pi*1.05)/(x+0.05))
mu.dop=dop.f(t)

@@

var.est.1=ashsmooth.gaus(X.s.1,v.est=TRUE)
var.est.2=ashsmooth.gaus(X.s.2,v.est=TRUE)
var.est.3=ashsmooth.gaus(X.s.3,v.est=TRUE)
var.est.4=ashsmooth.gaus(X.s.4,v.est=TRUE)

@@

```{r fig.width=12, fig.height=12}
par(mfrow=c(2,2))
plot(sigma.t.1^2, xlab = "", ylab = "",type='l', main = "V1")
lines(var.est.1,col=2)
plot(sigma.t.2^2, xlab = "", ylab = "",type='l', main = "V2")
lines(var.est.2,col=2)
plot(sigma.t.3^2, xlab = "", ylab = "",type='l', main = "V3")
lines(var.est.3,col=2)
plot(sigma.t.4^2, xlab = "", ylab = "",type='l', main = "V4")
lines(var.est.4,col=2)

@@

Now we take a look at how this wavelet-based approach handles unevenly spaced data. Wavelet methods are not optimized for unevenly spaced, but still works. We first look at the case when the data are normally spaced.

@@

##smash
mise(t(mu.est),mu.t,10)
##TI thresholding with variance estimated from smash
mise(t(mu.est.tivar.ash),mu.t,10)
##TI thresholding with variance estimated from running MAD
mise(t(mu.est.tivar.mad),mu.t,10)
##TI thresholding with constant variance (estimated)
mise(mu.est.ti,mu.t,10)
##EBayes with constant variance (estimated)
mise(mu.est.ti.ebayes,mu.t,10)

@@

mu.est=apply(X.s,1,ashsmooth.gaus)
mu.est.tivar.ash=apply(X.s,1,ti.thresh,method="smash")
mu.est.tivar.mad=apply(X.s,1,ti.thresh,method="rmad")

##smash
mise(t(mu.est),mu.t,10)
##TI thresholding with variance estimated from smash
mise(t(mu.est.tivar.ash),mu.t,10)
##TI thresholding with variance estimated from running MAD
mise(t(mu.est.tivar.mad),mu.t,10)
##TI thresholding with constant variance (estimated)
mise(mu.est.ti,mu.t,10)
##EBayes with constant variance (estimated)
mise(mu.est.ti.ebayes,mu.t,10)

```{r fig.width=12, fig.height=12}
par(mfrow=c(1,1))
plot(t,mu.t, xlab = "", ylab = "",type='l')
lines(t,mu.est[,1],col=2)
lines(t,mu.est.tivar.mad[,1],col=3)
lines(t,mu.est.ti[1,],col=4)
lines(t,mu.est.ti.ebayes[1,],col=6)
legend("topright", legend = c("smash", "ti_rmad", "ti_homo", "ebayes_homo"), col = c(2, 3, 4, 6))

Now for the case when the data are distributed according to a Poisson Process.

```{r}
n=1024
t=c(0,rexp(n-1,1))
t=cumsum(t)
t=(t-min(t))/(max(t)-min(t))

mu.est=apply(X.s,1,ashsmooth.gaus)
mu.est.tivar.ash=apply(X.s,1,ti.thresh,method="smash")
mu.est.tivar.mad=apply(X.s,1,ti.thresh,method="rmad")

##smash
mise(t(mu.est),mu.t,10)
##TI thresholding with variance estimated from smash
mise(t(mu.est.tivar.ash),mu.t,10)
##TI thresholding with variance estimated from running MAD
mise(t(mu.est.tivar.mad),mu.t,10)
##TI thresholding with constant variance (estimated)
mise(mu.est.ti,mu.t,10)
##EBayes with constant variance (estimated)
mise(mu.est.ti.ebayes,mu.t,10)

```{r fig.width=12, fig.height=12}
par(mfrow=c(1,1))
plot(t,mu.t, xlab = "", ylab = "",type='l')
lines(t,mu.est[,1],col=2)
lines(t,mu.est.tivar.mad[,1],col=3)
lines(t,mu.est.ti[1,],col=4)
lines(t,mu.est.ti.ebayes[1,],col=6)
legend("topright", legend = c("smash", "ti_rmad", "ti_homo", "ebayes_homo"), col = c(2, 3, 4, 6))

mu.est=apply(X.s,1,ashsmooth.gaus,ashparam=list(nullcheck=FALSE))
mu.est.tivar.ash=apply(X.s,1,ti.thresh,method="smash",ashparam=list(nullcheck=FALSE))
mu.est.tivar.mad=apply(X.s,1,ti.thresh,method="rmad")

##smash
mise(t(mu.est),mu.t,10)
##TI thresholding with variance estimated from smash
mise(t(mu.est.tivar.ash),mu.t,10)
##TI thresholding with variance estimated from running MAD
mise(t(mu.est.tivar.mad),mu.t,10)
##TI thresholding with constant variance (estimated)
mise(mu.est.ti,mu.t,10)
##EBayes with constant variance (estimated)
mise(mu.est.ti.ebayes,mu.t,10)

```{r fig.width=12, fig.height=12}
par(mfrow=c(1,1))
plot(t,mu.t, xlab  ="",  ylab = "",type='l')
lines(t,mu.est[,1],col=2)
lines(t,mu.est.tivar.mad[,1],col=3)
lines(t,mu.est.ti[1,],col=4)
lines(t,mu.est.ti.ebayes[1,],col=6)
legend("topright", legend = c("smash", "ti_rmad", "ti_homo", "ebayes_homo"), col = c(2, 3, 4, 6))

@@

  mu.est=ashsmooth.gaus(y.data)
  mu.est=mu.est[1:200]
  var.est=ashsmooth.gaus(y.data,v.est=TRUE)
  var.est=var.est[1:200]

@@

plot(x,mu.t, xlab = "", ylab = "",type='l', main = "mean function")
lines(x,mu.est[1:200],col=2)

plot(x,sigma.t^2, xlab = "", ylab = "",type='l', main = "var function")
lines(x,var.est[1:200],col=2)

@@

```{r}
data=read.csv("../code_paper/gaus_data/treas_bill/treas_bill.csv",header=FALSE)
data=as.numeric(data[,1])

@@

y.est=ashsmooth.gaus(y.final)
y.est.var=ashsmooth.gaus(y.final,v.est=TRUE)

@@

```{r fig.width=12, fig.height=12}
par(mfrow=c(2,2))
plot(data, xlab = "year", ylab = "interest rate",type='l')

plot(x,y)
lines(x.mod,y.est[1:(length(y.mod))], xlab = "X", ylab = "Y",col=2)

plot(x.mod,y.est[1:(length(y.mod))], xlab = "X", ylab = "Y",type='l',ylim=c(-0.3,0.3))

plot(x.mod,sqrt(y.est.var[1:(length(y.mod))]), xlab = "X",  ylab = "volatility",type='l')
````

@@

  mu.est<-ashsmooth.gaus(y.final)
  mu.est=mu.est[1:n]
  mse.hsm.c[i]=mse(mu.est,mu.t)

@@

  mu.est<-ashsmooth.gaus(y.final)
  mu.est=mu.est[1:n]
  mse.hsm.n[i]=mse(mu.est,mu.t)

@@

y.est=ashsmooth.gaus(y.final)
y.est=y.est[1:length(y)]
y.var.est=ashsmooth.gaus(y.final,v.est=TRUE)
y.var.est=y.var.est[1:length(y)]

@@

plot(x,y, xlab = "time", ylab = "acceleration",ylim=c(-150,100), main = "mean function")
lines(x,y.est,col=2)
plot(x,y.var.est, xlab = "time", ylab = "acceleration",type='l', main = "var function")

@@

plot(x,y, xlab = "ethonal", ylab = "NOx",ylim=c(-1,5), main = "mean function")
lines(x,y.est,col=2)
plot(x,y.var.est, xlab = "ethonal", ylab = "NOx",type='l', main = "var function")
